\documentclass[11pt]{article}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}

\def\N{{\mathbb N}}
\def\NN{{\mathcal N}}
\def\R{{\mathbb R}}
\def\E{{\mathbb E}}
\def\rank{{\mathrm{rank}}}
\def\tr{{\mathrm{trace}}}
\def\P{{\mathrm{Prob}}}
\def\sign{{\mathrm{sign}}}
\def\diag{{\mathrm{diag}}}

\setlength{\oddsidemargin}{0.25 in}
\setlength{\evensidemargin}{-0.25 in}
\setlength{\topmargin}{-0.6 in}
\setlength{\textwidth}{6.5 in}
\setlength{\textheight}{8.5 in}
\setlength{\headsep}{0.75 in}
\setlength{\parindent}{0.25 in}
\setlength{\parskip}{0.1 in}

\newcommand{\lecture}[4]{
   \pagestyle{myheadings}
   \thispagestyle{plain}
   \newpage
   \setcounter{page}{1}
   \setcounter{section}{0}
   \noindent
   \begin{center}
   \framebox{
      \vbox{\vspace{2mm}
    \hbox to 6.28in { {\bf A Mathematical Introduction to Data Science \hfill #4} }
       \vspace{6mm}
       \hbox to 6.28in { {\Large \hfill #1  \hfill} }
       \vspace{6mm}
       \hbox to 6.28in { {\it Instructor: #2\hfill #3} }
      \vspace{2mm}}
   }
   \end{center}
   \markboth{#1}{#1}
   \vspace*{4mm}
}


\begin{document}

\lecture{Mini-Project 1}{Yuan Yao}{Due: October 6?, 2017}{September 29, 2017}

%The problem below marked by $^*$ is optional with bonus credits. % For the experimental problem, include the source codes which are runnable under standard settings. 
%
%\begin{enumerate}
%
%\item {\em Manifold Learning}: The following codes by Todd Wittman contain major manifold learning algorithms talked on class.
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/Spring2011/matlab/mani.m}
%
%Precisely, eight algorithms are implemented in the codes: MDS, PCA, ISOMAP, LLE, Hessian Eigenmap, Laplacian Eigenmap, Diffusion Map, and LTSA. 
%The following nine examples are given to compare these methods,
%\begin{enumerate}
%\item Swiss roll;
%\item Swiss hole;
%\item Corner Planes;
%\item Punctured Sphere;
%\item Twin Peaks;
%\item 3D Clusters;
%\item Toroidal Helix;
%\item Gaussian;
%\item Occluded Disks.
%\end{enumerate}
%Run the codes for each of the nine examples, and analyze the phenomena you observed. 
%
%\end{enumerate}

%\newpage


\section{Mini-Project Requirement and Datasets}

This project aims to exercise the tools in the class, such as random projections, robust PCA, sparse PCA, etc., based on the real datasets. In the below, we list some candidate datasets for your reference. 

\begin{enumerate}
\item Pick up ONE (or more if you like) favorite dataset below to work. If you would like to work on a different problem outside the candidates we proposed, please email course instructor about your proposal.  
\item Team work: we encourage you to form small team, up to THREE persons per group, to work on the same problem. Each team just submit ONE \emph{poster} report, \emph{with a clear remark on each person's contribution}. A sample poster file with PKU logo can be found at \\ 
\url{http://math.stanford.edu/~yuany/course/reference/poster_v5.pdf} \\
whose source LATEX codes can be downloaded at \\
\url{http://math.stanford.edu/~yuany/course/reference/pkuposter.zip} 
\item In the report, (1) design or raise your scientific problems (a good problem is sometimes more important than solving it); (2) show your results with your careful analysis of the results toward answering your problem. Remember: scientific analysis and reasoning are more important than merely the performance results. Source codes may be submitted through email as a zip file, or as an appendix if it is not large.    
\item Submit your report by email or paper version no later than the deadline, to Teaching Assistant (TA), Huizhuo Yuan (\href{mailto:datascience\_hw@126.com}{datascience\_hw@126.com}). 
\end{enumerate}

\section{Co-appearance data in novels: Dream of Red Mansion and Journey to the West}

A 374-by-475 binary matrix of character-event can be found at the course website, in .XLS, .CSV, .RData, and .MAT formats. For example the RData format is found at

\url{http://math.stanford.edu/~yuany/course/data/dream.RData} 

\noindent with a readme file:

\url{http://math.stanford.edu/~yuany/course/data/dream.Rd}

\noindent as well as the .txt file which is readable by R command {\tt read.table()},

\url{http://math.stanford.edu/~yuany/course/data/HongLouMeng374.txt}

\url{http://math.stanford.edu/~yuany/course/data/readme.m}

Thanks to Ms. WAN, Mengting, who helps clean the data and kindly shares her BS thesis for your reference
 
\url{http://math.stanford.edu/~yuany/report/WANMengTing2013_HLM.pdf}

%Among various choices of analysis, with this data matrix $X$, you may form a weighted graph $W=X * X'$, pursue PCA of $X$, and sparse SVD of $X$ etc. As an example, here is a project presentation by LI, Liying which gives an analysis of A Journey to the West (by Chen-En Wu) based on PCA, for the class Mathematical Introduction to Data Science in Fall 2012 where you may find more interesting approaches.
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/reference/LiyingLI_Xiyouji2012_slides.pdf}

Moreover you may find a similar matrix of 302-by-408 for the Journey to the West (by Chen-En Wu) at:

\url{http://math.stanford.edu/~yuany/course/data/west.RData}

\noindent whose matlab format is saved at

\url{http://math.stanford.edu/~yuany/course/data/xiyouji.mat}

%%%%%%

\section{Jiashun Jin's data on Coauthorship and Citation Networks for Statisticians}
Thanks to Prof. Jiashun Jin at CMU, who provides his collection of citation and coauthor data for statisticians. The data set covers all papers between 2003 and the first quarter of 2012 from the Annals of Statistics, Journal of the American Statistical Association, Biometrika and Journal of the Royal Statistical Society Series B. The paper corrections and errata are not included. There are 3607 authors and 3248 papers in total. The zipped data file (14M) can be found at 

\url{http://math.stanford.edu/~yuany/course/data/jiashun/Jiashun.zip}

\noindent with an explanation file

\url{http://math.stanford.edu/~yuany/course/data/jiashun/ReadMe.txt}

With the aid of Mr. LI, Xiao, a subset consisting 35 COPSS award winners (\url{https://en.wikipedia.org/wiki/COPSS_Presidents\%27_Award}) up to 2015, is contained in the following file

\url{http://math.stanford.edu/~yuany/course/data/copss.txt} 

\noindent An example was given in the following article, A Tutorial of Libra: R Package of Linearized Bregman Algorithms in High Dimensional Statistics, downloaded at

\url{http://math.stanford.edu/~yuany/course/reference/Libra_Tutorial_springer.pdf}

The citation of this dataset is: \emph{P. Ji and J. Jin. Coauthorship and citation networks for statisticians. arXiv:1410.2840, 2014}. As the paper has not been formally published yet, please do not use the dataset outside this class or for any kinds of publications without the permission of the authors. 

\section{Drug Efficacy Data}

Thanks to Prof. Xianting Ding at Shanghai Jiao Tong University and Prof. Chih-Ming Ho from University of California at Los Angeles, we have the following datasets on combinatorial drug efficacy.

The first dataset consists of two experiments, all with the same 4 drugs in cell lines for attacking leukemia, with 256 experiments of combinatorial drug dosage at 4 levels. The response is the therapeutic window measuring the efficacy with a trade-off by toxicity. 
 
 \url{http://math.stanford.edu/~yuany/course/data/Ding_4drugs.xlsx}

\noindent whose drugs are explained in 

\url{http://math.stanford.edu/~yuany/course/data/Ding_4drugs_readme.pdf}

Can you find a good prediction of drug response efficacy using those combinatorial dosage levels? It was suggested that quadratic polynomials at logarithmic dosage levels are good models in personalized medicine, e.g. the following cover paper in Science \emph{Translation Medicine}:

\url{http://stm.sciencemag.org/content/8/333/333ra49}

\noindent with a sample 14 drug efficacy at level 2 experiment data in liver transplant: 

\url{http://math.stanford.edu/yuany/course/data/TB-FSC-03A-data.xlsx}

%\subsection{The Characters in A Dream of Red Mansion} 
%
%A 376-by-475 matrix of character-event can be found at the course website, in .XLS, .CSV, and .MAT formats. For example the Matlab format is found at
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/hongloumeng/hongloumeng376.mat} 
%
%\noindent with a readme file:
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/hongloumeng/readme.m}
%
%Thanks to Ms. WAN, Mengting (now at UIUC), an update of data matrix consisting 374 characters (two of 376 are repeated) which is readable by R read.table() can be found at 
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/hongloumeng/HongLouMeng374.txt}
%
%\noindent She also kindly shares her BS thesis for your reference
% 
% \url{http://www.math.pku.edu.cn/teachers/yaoy/reference/WANMengTing2013_HLM.pdf}
%
%% Among various choices of analysis, with this data matrix $X$, you may form a weighted graph $W=X * X'$, pursue PCA of $X$. 
%
%\subsection{A Journal to the West} On course website, you may also find the link to this dataset with a 302-by-408 matrix, whose matlab format is saved at
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/Fall2011/xiyouji/xiyouji.mat}
%
%For your reference, here is a project presentation by Mr. LI, Liying (at PKU) which gives an analysis based on PCA
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/reference/LiyingLI_Xiyouji2012_slides.pdf}
%

\section{Heart PCI Operation Effect Prediction}

The following data, provided by Dr. Jinwen Wang at Anzhen Hospital, 

\url{http://math.stanford.edu/~yuany/course/data/heartData_20140401.xlsx}

\noindent contains 2581 patients with 73 measurements (inputs) as well as a response variable indicating if after the heart operation there is a null-reflux state. This is a classification problem, with a challenge from the large amount of missing values. Sheet 3 and 4 in the file contains some explanation of the data and variables. 

The problems are listed here:
\begin{enumerate}
\item The inputs (covariates) are of three kinds, measurements upon check-in, measurements before PCI operation, and measurements in PCI operations. For doctors, it is desired to find a prediction model based on measurements before the operation (including check-in). Sheet 2 in the file contains only such measurements.
\subitem The following two reports by LV, Yuan and LI, Xiao, respectively, might be interesting to you:

\url{http://math.stanford.edu/~yuany/course/reference/MSThesis.LvYuan.pdf} 

\url{http://arxiv.org/abs/1511.04656} 

\item It is also an interesting problem how to predict the effect based on all measurements, with lots of missing values. Sheet 1 contains the full measurements. There are some good work by previous students, which are listed here for your reference: 
%\subitem The following two reports by LU, Yu and WANG, Qing, are probably inspiring to you.
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/reference/LuYu_201303_BigHeart.pdf} 
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/reference/WangQing_201303_BigHeart.pdf} 

\subitem The following report by MIAO, Wang and LI, Yanfang, pioneers in missing value treatment. 

\url{http://math.stanford.edu/~yuany/course/reference/MiaoLi2013S_project01.pdf}

\end{enumerate} 

%\emph{In the final project, it is desired to take only those measurements upon check-in to predict the probability of non-reflux (non-reflow) after PCI operations. An interpretable model adds a big value! You may compare with your first warm-up project to show your improvements.} 

\section{Identification of Raphael's paintings from the forgeries}

The following data, provided by Prof. Yang WANG from HKUST,

\url{https://drive.google.com/folderview?id=0B-yDtwSjhaSCZ2FqN3AxQ3NJNTA&usp=sharing}

\noindent contains a 28 digital paintings of Raphael or forgeries. Note that there are both jpeg and tiff files, so be careful with the bit depth in digitization. The following file

\url{https://docs.google.com/document/d/1tMaaSIrYwNFZZ2cEJdx1DfFscIfERd5Dp2U7K1ekjTI/edit}

\noindent contains the labels of such paintings, which are 
\begin{enumerate}
\item[1] Maybe Raphael - Disputed
\item[2] Raphael
\item[3] Raphael
\item[4] Raphael
\item[5] Raphael
\item[6] Raphael
\item[7] Maybe Raphael - Disputed
\item[8] Raphael
\item[9] Raphael
\item[10] Maybe Raphael - Disputed
\item[11] Not Raphael
\item[12] Not Raphael
\item[13] Not Raphael
\item[14] Not Raphael
\item[15] Not Raphael
\item[16] Not Raphael
\item[17] Not Raphael
\item[18] Not Raphael
\item[19] Not Raphael
\item[20] My Drawing (Raphael?)
\item[21] Raphael
\item[22] Raphael
\item[23] Maybe Raphael - Disputed
\item[24] Raphael
\item[25] Maybe Raphael - Disputed
\item[26] Maybe Raphael - Disputed
\item[27] Raphael
\item[28] Raphael
\end{enumerate}
Can you exploit the known Raphael vs. Not Raphael data to predict the identity of those 6 disputed paintings (maybe Raphael)? The following student poster report seems a good exploration

\url{http://math.stanford.edu/~yuany/course/2015.fall/poster/Raphael_LI\%2CYue_1300010601.pdf}

The following paper by Haixia Liu, Raymond Chan, and me studies Van Gogh's paintings which might be a reference for you:

\url{http://dx.doi.org/10.1016/j.acha.2015.11.005}

\section{Finance Data}
The following data contains 1258-by-452 matrix with closed prices of 452 stocks in SNP'500 for workdays in 4 years.

\url{http://www.math.pku.edu.cn/teachers/yaoy/data/snp452-data.mat}

%You may use PCA to explore the `invisible hands' of markets.

\section{Hand-written Digits} The website 

\url{http://www-stat.stanford.edu/\~tibs/ElemStatLearn/datasets/zip.digits/}

contains images of 10 handwritten digits (`$0$',...,`9');


%\section{Air Quality Weibo Data} (courtesy of Prof. Xiaojin Zhu from University of Wisconsin at Madison) 
%You can login my server:
%
%\texttt{ssh einstein@162.105.205.92}
%
%\noindent using the password I provided on class. 
%
%On the read-only folder \texttt{/data/AQweibo/}, the \texttt{AQICityData/} directory contains the Weibo posts, the AQI for 108 cities with (AQI) information during the study period
%from 2013-11-18 to 2013-12-18 (both inclusive); Information for the spatiotempral bin (city,date) is in the directory \texttt{city\_date/}. See \texttt{README.txt} for more information.
%
%

\section{SNPs Data}
 This dataset contains a data matrix $X\in \R^{p\times n}$ of about $n=650,000$ columns of SNPs (Single Nucleid Polymorphisms) and $p=1064$ rows of peoples around the world. Each element is of three choices, $0$ (for `AA'), $1$ (for `AC'), $2$ (for `CC'), and some missing values marked by $9$. 

\url{http://math.stanford.edu/~yuany/course/ceph_hgdp_minor_code_XNA.txt.zip}

\noindent which is big (151MB in zip and 2GB original txt). Moreover, the following file contains the region where each people comes from, as well as two variables {\texttt{ind1}} and{\texttt{ind2}} such that $X({\texttt{ind1}},{\texttt{ind2}})$ removes all missing values. 

\url{http://math.stanford.edu/~yuany/course/data/HGDP_region.mat}

Some results by PCA can be found in the following paper, Supplementary Information. 

\url{http://www.sciencemag.org/content/319/5866/1100.abstract}

%Attention: this last dataset is relatively big with about 2GB size. 

You can login my server:

\texttt{ssh einstein@162.105.205.92}

\noindent using the password I provided on class. On the read only folder \texttt{/data/snp/}, you will find all the data in both .txt and .mat (\texttt{data.mat, HGDP\_region.mat, readme.m}).



%\subsection{Bird Flu Dataset} (courtesy of Steve Smale and Cissy) This dataset 162 H5N1 (bird flu) virus sequences discovered around the world:
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_seq162.txt} 
%
%Locations of such virus discovered are reported with latitude and longitude coordinates on the globe:
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_latgrat.txt} 
%
%Pairwise geodesic distances between these 162 sites are constructed as  
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_geodist.txt}
%
%A kernel-induced $l_2$-distances between 162 virus sequences are given in 
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_l2dist.txt}
\end{document}


