---
title: "Untitled"
author: "Esther Visser"
date: "28 november 2017"
output: html_document
---


```{r setup, include=FALSE}
setwd("~/Universiteit/Master/Mathematical_introduction_to_data_science")
```

```{r}
list<-array(read.table("pairwisevotingsproject2.txt",sep=","))
winset<-cbind(list[2:79,4],list[2:79,6])
test<-c()
for(i in 1:78){test<-append(test,(as.numeric(winset[i,1])))}
for(i in 1:78){test<-append(test,(as.numeric(winset[i,2])))}
test<-unique(sort(test))
```
So basically every winner id corresponds with a number 1 up to 12.
```{r}
c(winset[1,1])
c(winset[2,1])
c(winset[5,1])
c(winset[6,1])
c(winset[8,1])
c(winset[9,1])
c(winset[11,1])
c(winset[1,2])
c(winset[2,2])
c(winset[5,2])
c(winset[6,2])
c(winset[17,2])

```
In this list we see all the numbers occur once, hence we can see what they are in the original list. We see that the numbers we get correspond to how they are called, so for example for number 11 we see that list[2,5] gives 11. Siao Li.

First we get a ranking using Borda count: we give every winner 1 point, every loser 0. Then we divide it by the number of times they have been an option to vote for (if there are enough votings this will make no difference, however on a small amount of votings it might give a huge difference if your name popped up more often).
```{r}
win<-c()
for(i in 1:78){win<-append(win,(as.numeric(winset[i,1])))}
win<-sort(win)
lose<-c()
for(i in 1:78){lose<-append(lose,(as.numeric(winset[i,2])))}
lose<-sort(lose)
score<-function(n){
  wintimes<-length(win[win==n])
  losetimes<-length(lose[lose==n])
  return(wintimes/(wintimes+losetimes))
}
```


```{r}
borda<-data.frame(winner=c(1:12),score=sapply(c(1:12),score))
borda[with(borda,order(score,decreasing = TRUE)),]
```

Second way of getting a ranking: kennemy optimization: minimize the number of mismatches, however, this is NP hard, so must run for a long time.
```{r}
mismatch<-function(ord){
  count<-0
  for(i in 1:78){
    val1<-as.numeric(winset[i,1])
    val2<-as.numeric(winset[i,2])
    place1<-which(ord==val1)
    place2<-which(ord==val2)
    if(place1>place2){
      count<-count+1
    }
  }
  return(count)
}
```

```{r}
mismatch(c(1,2,3,4,5,6,7,8,9,10,11,12))
mismatch(c(10,3,4,2,11,5,6,8,9,7,1,12))
mismatch(c(10,3,2,4,11,5,6,8,9,7,1,12))
```
```{r}
for(bb in 1:40){
set.seed(bb)
x<-sample(1:12,12,replace=FALSE)
scorex<-mismatch(x)
bestscore<-scorex
bestperm<-x
repeat{
  x<-bestperm
  scorex<-mismatch(x)
for(i in 1:12){for(j in 1:12){
  xperm<-x
  xperm[c(i,j)] <- x[c(j,i)]
  scoreperm<-mismatch(xperm)
  if(scoreperm<bestscore){
    bestperm<-xperm
    bestscore<-scoreperm
  }
}}
if(bestscore == scorex){break}
}
print(c(bestscore,bb))}
```

We just locally minimize the score and see the best value
set.seed(1) gives 16, with permutation 10  3  2  5  4  8  6 11  9  1  7 12
2 gives bestscore 17, 3 gives 18.
4 gives 15 with permutation 10 11  3  2  4  5  8  6  1  7 12  9
20,37 and 40 also give 15

Now we can perform the same algorithm, only interchanging 4 instead of 2 elements
```{r}
bestscoresofar<-15
numberseed<-0
for(bb in 21:40){
set.seed(bb)
x<-sample(1:12,12,replace=FALSE)
scorex<-mismatch(x)
bestscore<-scorex
bestperm<-x
repeat{
  x<-bestperm
  scorex<-mismatch(x)
for(i1 in 1:12){for(j1 in i1:12){for(i2 in i1:12){for(j2 in i2:12){
  xperm<-x
  xperm[c(i1,j1)] <- x[c(j1,i1)]
  xperm[c(i2,j2)] <- xperm[c(j2,i2)]
  scoreperm<-mismatch(xperm)
  if(scoreperm<bestscore){
    bestperm<-xperm
    bestscore<-scoreperm
  }
}}}}
if(bestscore == scorex){break}
}
if(bestscore<=bestscoresofar){
  print(c(bb,bestscore,bestperm))
  bestscoresofar<-bestscore
  }
}
```

For now optimal sets
[1]  2 15 10 11  3  2  4  5  8  6  1  7 12  9
[1]  4 15 10 11  3  2  4  5  8  6  1  7 12  9
[1]  6 15 10 11  3  2  5  4  8  9  6  1  7 12
[1]  9 15 10 11  3  2  4  5  8  9  6  1  7 12
[1] 20 15 10 11  3  2  4  5  8  6  1  7  9 12
[1] 27 15 10 11  3  5  2  4  8  6  1  7  9 12
[1] 28 15 10 11  3  2  5  4  9  8  6  1  7 12
[1] 32 15 10 11  3  5  2  4  9  8  6  1  7 12
[1] 33 15 10 11  3  2  5  4  9  8  6  1  7 12
[1] 34 15 10 11  3  5  2  4  8  6  1  9  7 12
[1] 35 15 10 11  3  5  2  4  8  6  1  9  7 12
[1] 37 15 10 11  3  5  2  4  8  6  1  9  7 12
[1] 40 15 10 11  3  2  5  4  8  6  1  7 12  9
 As one can see it converges to almost the same set everytime, so it looks like that the global minimum has 15 mismatches.
 
 
```{r}
c(10,11,3,2,4,5,8,6,1,7,12,9) #2,4
c(10,11,3,2,5,4,8,9,6,1,7,12) #6, 9
c(10,11,3,2,5,4,9,8,6,1,7,12) #28,32,33
```


# I don't use the following as it took too long to calculate all permutations:
```{r}
library(gtools)
```
Note that the function permutations(12,12) gives all permutations for the possible set. Now we can apply the function mismatch to that one

```{r}
perm<-permutations(12,12)
n<-dim(perm)[1]
bestscore<-100
bestperm<-c(1)
for(i in 1:n){
  scor<-mismatch(perm[i,])
  if(scor<bestscore){
    bestperm<-perm[i,]}
  }
```
Then find the best permutation
```{r}
bestperm
```


Third way of ranking:
  Define a random walk on (1,...,12) where we go from node i to note j with probability P_{ij}=1/dmax x a_{ij}/(a_{ji}+a_{ij}), where a_{ij} is the number of times that i wins from j. dmax is defined to be the maximum outdegree of a node
https://arxiv.org/pdf/1209.1688v2.pdf
http://web.engr.illinois.edu/~swoh/slide_ismp2012.pdf
In this case dmax=11. Then with probability 1-\sum_{j\neq i}P_{ij} the walk stays on the same node
```{r}
a<-function(i,j){
  val<-0
  for(n in 1:78){
    if(as.numeric(winset[n,1])==j&&as.numeric(winset[n,2])==i){
      val<-val+1
    }
  }
  return(val)
}
```
now define the probability of going to another node.
```{r}
p<-function(i,j){
  if(i!=j){if(a(i,j)+a(j,i)>0){
   return(1/11 * a(i,j)/(a(i,j)+a(j,i)))
  } else{return(0)}
  } else{
    return(1-sum(sapply(c(1:12)[c(1:12)!=i],function(j) if(a(i,j)+a(j,i)!=0){1/11 * a(i,j)/(a(i,j)+a(j,i))} else{0})))
}
}
mat<-matrix(0,nrow=12,ncol=12)
for(i in 1:12){for(j in 1:12){mat[i,j]<-p(i,j)}}
```
Note that the definition for P makes sense, as if j never wins of i, then we want the probability of of going from node i to note j (i.e. P(i,j)) to be 0, and indeed we defined a(i,j)=0 in this case, as a(i,j) counts the number of times j wins of i.


```{r}
recursive.rwalk=function(n){
  if (n==0) return(1)
  rw <- recursive.rwalk(n-1)
  return(c(rw,sample(1:12,1,prob=as.vector(mat[rw[length(rw)],]),replace=TRUE)))
}
set.seed(123)
vect<-recursive.rwalk(1000)
```
Note that the set.seed is a function determining the randomwalk. Basically this function makes the result repreducible, yet still (pseudo)random.
Now the random walk will be the most time on the nodes which are the best. So we get again a ranking
```{r}
scorerandomwalk<-function(n){
  return(length(vect[vect==n]))
}
rwalk<-data.frame(winner=c(1:12),score=sapply(c(1:12),scorerandomwalk))
rwalk[with(rwalk,order(score,decreasing = TRUE)),]
```

Fourth method: Condorcet method
Note that this method is mainly developped to find the winner, and not the order of all the other ones. According to http://civs.cs.cornell.edu/rp.html there are multiple methods, all of which give the same result if there is a Condorcetwinner

For the Concordetmethod, we must find out if one poster beats all other posters. Suppose poster i is the winner. This means that i beats every other poster j more than j beats i. As a(i,j) denotes the number of times that j wins of i, we can easily find this
```{r}
for(i in 1:12){
print(sapply(c(1:12)[c(1:12)!=i],function(j) a(j,i)-a(i,j)))
}
```
As one can see, 10 has a tie with everyone or wins, so 10 is the concordetwinner.





