---
title: "Karate club"
author: "Esther Visser"
date: "3 november 2017"
output: html_document
---
```{r}
setwd("~/Universiteit/Master/Mathematical_introduction_to_data_science")
```

```{r}
library(R.matlab)
mat<-readMat("karate.mat")
table<-mat$A
```

```{r}
d<-function(n){rowSums(table)[n]}
D<- diag(d(1:34))
Dinv<-diag(1/d(1:34))
P<- Dinv %*% table
```
As rowSums(P)=1 for all rows, this defines a chain, by just, if you are in a point i, then you take the ith row and you go to person j with probability P(i,j)
stationary distribution:
```{r}
pi<-c(d(1:34))
as.vector(t(P)%*% pi)==pi
```
Now part 3, see notes page 123. First we define the laplacion and the second eigenvector.
```{r}
lap<-matrix(nrow=34,ncol=34)
for(i in 1:34){
  for(j in 1:34){
    if(i==j){
      lap[i,j]<-1
    } else{
    if(table[i,j]!=0){
      lap[i,j]<--1/sqrt(d(i)*d(j))
    } else{
      lap[i,j]<-0
    }}
  }}
eigen<-eigen(lap)$vectors[,2]
sqrt(solve(D))%*%eigen
```

Now we find the cut by solving the equation in the lecturenotes (which we have to make lowerdimensional as 2 values of q are already given)
```{r}
M<-diag(34)-P
q<-solve(M[2:33,2:33],-M[2:33,34])
q<-append(append(0,q),1)
vplus<-c()
vmin<-c()
for(i in 1:34){
  if(q[i]<=0.5){vmin<-append(vmin,i)} else{vplus<-append(vplus,i)}
}
```

```{r}
library(qgraph)
```
```{r}
qgraph(table,groups=list(data1,data2))
qgraph(table, groups=list(vmin,vplus))
```
Now we define the functions J and Jplus
```{r}
J<-function(x,y){
  if(x!=y){
  pi[x]*(1-q[x])*P[x,y]*q[y]} else{0}
}
Jplus<-function(x,y){
  max(J(x,y)-J(y,x),0)
}
```
transitionflux
```{r}
T<-function(x){
  if(x==1){sum<-0
  for(i in 1:34){sum<-sum+Jplus(x,i)
  return(sum)}} else{
    sum<-0
  for(i in 1:34){sum<-sum+Jplus(i,x)
  return(sum)}
  }
}
```
The next two vectors are just the functions T and Jplus as a vector respectively matrix to be able to give weights in the directed graph
```{r}
vectT<-c()
for(i in 1:34){vectT<-append(vectT,T(i))}
matJ<-matrix(nrow=34,ncol=34)
for(i in 1:34){for(j in 1:34){matJ[i,j]<-Jplus(i,j)}}
```
Remove the values in R which are zero but keep the arrows if there is no flow, else the graph gets disconnected:
```{r}
for(i in 1:34){for(j in i:34){
  if(table[i,j]==1){
    if(Jplus(i,j)>0){table[j,i]<-NA}
    if(Jplus(j,i)>0){table[i,j]<-NA}
  }
}}
```
plot the graph
```{r}
qgraph(table+matJ, groups=list(vmin,vplus),vsize=4*(vectT+1),esize=10)
```

